Enhanced Global PTM Discovery (G-PTM-D) with MetaMorpheus

### 2018-02-15 paper.tex
djfkf
review
okay

\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
\usepackage[T1]{fontenc}       % Use modern font encodings
\usepackage{multirow}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{hyperref}

@@

\author{Stefan K. Solntsev}
\author{Michael R. Shortreed}
\author{Brian L. Frey}
\author{Lloyd M. Smith}
\email{smith@chem.wisc.edu}
\affiliation[UwMadison]
{University of Wisconsin-Madison}

@@

\begin{figure}[H]
  \includegraphics[scale=0.5]{paperDiagram.png}
  \caption{The three steps of the G-PTM-D workflow are indicated in the light beige color. The dark orange color corresponds to the inputs.}
  \label{fgr:diagram}
\end{figure}

@@

The database augmentation step allowed 208 modification types, including PTMs, adducts, and chemical modifications.
See Table S-1 for a complete listing.

@@

We propose a limited multi-notch search as a remedy to the trade-offs caused by choosing a narrow (few ppm) or wide (few Da) precursor mass tolerance during a traditional narrow-window search.
This limited multi-notch search allows mass differences between precursor and theoretical peptides of 0, 1.0029 and 2.0054 Da, but it limits the width around these three mass differences to a few ppm (e.g. 1-20 ppm).
This approach provides a means of getting an increased number of correct identifications while mitigating problems with incorrect identifications.
This alternative would appear to result in loss of identification of co-isolated peptides, but MetaMorpheus is programmed to handle this problem by identifying multiple co-isolated precursor masses (see section below).

@@

@@

Further improvements arise from using a multi-notch search to create the augmented database.
The enhanced G-PTM-D procedure employed a narrow 5 ppm tolerance around 208 mass difference notches (see Document S-2).
The resulting augmented database is greatly improved compared to using an open search for this step.
At this point, another advantage of the enhanced approach is clear; the overall search time dropped significantly from 22.1 hours to 2.9 hours, primarily due to the narrow tolerance multi-notch search versus the open search.

@@

%\end{suppinfo}

>>>>

\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
\usepackage[T1]{fontenc}       % Use modern font encodings
\usepackage{multirow}
\usepackage{subcaption}
\usepackage{booktabs}

@@

\author{Stefan K. Solntsev}
\author{Michael R. Shortreed}
\author{Brian L. Frey}
\author{Lloyd M. Smith}
\email{smith@chem.wisc.edu}
\altaffiliation{Genome Center of Wisconsin, University of Wisconsin-Madison, Madison, WI, USA}
\affiliation[UW Madison]
{Department of Chemistry, University of Wisconsin-Madison, Madison, WI, USA}

@@

\begin{figure}[H]
  \includegraphics[scale=0.75]{paperDiagram.png}
  \caption{Comparing the original G-PTM-D workflow with the proposed enhanced workflow. Major changes are highlighted in pink.}
  \label{fgr:diagram}
\end{figure}

@@

The database augmentation step allowed 208 modification types, including PTMs, adducts, and chemical modifications.
See supplement Supp1 for a complete listing.

@@

A multi-notch search is a remedy to the trade-offs caused by choosing a narrow (few ppm) or wide (few Da) precursor mass tolerance during a traditional narrow-window search.
It is not novel, and many useful search tools such as Tandem, MSGF+, Coment, MSFragger, Sequest have this search functionality under names such as "isotope error option".
This limited multi-notch search allows mass differences between precursor and theoretical peptides of 0, 1.0029 and 2.0054 Da, but it limits the width around these three mass differences to a few ppm (e.g. 1-20 ppm).
This approach provides a means of getting an increased number of correct identifications while mitigating problems with incorrect identifications.
This alternative would appear to result in loss of identification of co-isolated peptides, but MetaMorpheus is programmed to handle this problem by identifying multiple co-isolated precursor masses (see section below).

@@

The results of a multi-notch search are assessed using notch-specific FDRs, computed individually for each notch, similarly to MsFragger.

@@

Further improvements arise from using a multi-notch search to create the augmented database.
The enhanced G-PTM-D procedure employed a narrow 5 ppm tolerance around 208 mass difference notches (see supplement Supp2Calibration).
The resulting augmented database is greatly improved compared to using an open search for this step.
At this point, another advantage of the enhanced approach is clear; the overall search time dropped significantly from 22.1 hours to 2.9 hours, primarily due to the narrow tolerance multi-notch search versus the open search.

@@

\end{suppinfo}
